---
title: "Final Project"
author: "Alexander Hopkins, Brett Hunsanger & Ted Dove"
output: 
  html_document:
    keep_md: TRUE
    toc: TRUE
    toc_float: TRUE
    df_print: paged
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error=TRUE, message=FALSE, warning=FALSE)
```

```{r libraries}
library(tidyverse)     # for data cleaning and plotting
library(lubridate)     # for date manipulation
library(openintro)     # for the abbr2state() function
library(maps)          # for map data
library(ggmap)         # for mapping points on maps
library(gplots)        # for col2hex() function
library(RColorBrewer)  # for color palettes
library(sf)            # for working with spatial data
library(leaflet)       # for highly customizable mapping
library(ggthemes)      # for more themes (including theme_map())
library(plotly)        # for the ggplotly() - basic interactivity
library(gganimate)     # for adding animation layers to ggplots
library(gifski)        # for creating the gif (don't need to load this library every time,but need it installed)
library(transformr)    # for "tweening" (gganimate)
library(shiny)         # for creating interactive apps
library(patchwork)     # for nicely combining ggplot2 graphs  
library(gt)            # for creating nice tables
library(rvest)         # for scraping data
library(robotstxt)     # for checking if you can scrape data
library(readxl)
library(scales)
theme_set(theme_minimal())
```

```{r data}
#Eviction data from the Eviction Lab
evictions_state <- read_csv("evictions_state.csv")
states_midpoint <- read_csv("state-midpoints.csv")

states_map <- map_data("state")
```

### The Data
Our data is from the [Eviction Lab](https://evictionlab.org/), on their website they introduce themselves as "The Eviction Lab is a team of researchers, students, and website architects who believe that a stable, affordable home is central to human flourishing and economic mobility. Accordingly, understanding the sudden, traumatic loss of home through eviction is foundational to understanding poverty in America. Drawing on tens of millions of records, the Eviction Lab at Princeton University has published the first ever dataset of evictions in America, going back to 2000."

### Our Initial Research Questions
When we began thinking about the data we had some questions right away about eviction and race. Is there a relationship between race and the eviction rate in the US? There is also rent information in the data and we wanted to compare rent over each state and see what effect rent had on eviction rates. Some of the final questions we had were what are the states with the highest eviction filing rates and if it was easier to evict people in some states compared to others.

### Some Preliminary Graphs
First we wanted to explore the data with a plotly graph and mapped the eviction rates by state in 2016. After looking at that graph and discussing more what we wanted to show with our data (Courtesy of Eviction Lab) we decided to make a shiny app that users could interact with to see eviction rates over time as well as interact with demographic and other data by state to see relationships between that data and eviction rates. 

```{r}
evict_rate_map <- evictions_state %>% 
  filter(year == 2016) %>% 
  mutate(name = str_to_lower(name)) %>% 
  ggplot() +
  geom_map(map = states_map,
           aes(map_id = name,
               fill = `eviction-rate`)) +
  expand_limits(x = states_map$long, y = states_map$lat) + 
  theme_map() +
  labs(title = "",
       fill = "") +
  scale_fill_viridis_c() +
  theme(legend.background = element_blank()) +
  labs(title = "Eviction Rate by State in 2016")

  ggplotly(evict_rate_map)
```

This graph for example overlays evictions rates on a map of the US that is colored by population. We can see that population does not seem to have a strong effect on eviction rates as California and Texas do not have greatly high eviction rates compared to some states like South Carolina and Delaware which have much lower populations.

```{r}
options(scipen = 9)
evictions_state %>% 
  left_join(states_midpoint,
            by = c("name" = "location")) %>% 
  filter(year == 2016,
         name != "Hawaii") %>% 
  mutate(name = str_to_lower(name)) %>% 
  ggplot() +
  geom_map(map = states_map,
           aes(map_id = name,
               fill = population)) +
  geom_point(aes(x = lon, y = lat, size = `eviction-rate`),
             color = "red") +
  expand_limits(x = states_map$long, y = states_map$lat) + 
  theme_map() +
  labs(title = "",
       fill = "Population",
       size = "Eviction Rate") +
  scale_fill_viridis_c(labels = comma) +
  theme(legend.background = element_blank(),
        legend.position = "right") +
  labs(title = "Population and Eviction rates in 2016")
```


### Our Shiny App
Click [Here]() to interact with the data yourself! Choose from a variety of factors like race, median rent, and year to display graphs to see relationships between demographics and evictions. The first tab displays statewide data and the second "by state" tab displays individual states and their county level data.

### Final Remarks


**Include a few more observations here**

