---
title: "Final Project"
author: "Alexander Hopkins, Brett Hunsanger & Ted Dove"
output: 
  html_document:
    keep_md: TRUE
    toc: TRUE
    toc_float: TRUE
    df_print: paged
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error=TRUE, message=FALSE, warning=FALSE)
```

```{r libraries}
library(tidyverse)     # for data cleaning and plotting
library(lubridate)     # for date manipulation
library(openintro)     # for the abbr2state() function
library(maps)          # for map data
library(ggmap)         # for mapping points on maps
library(gplots)        # for col2hex() function
library(RColorBrewer)  # for color palettes
library(sf)            # for working with spatial data
library(leaflet)       # for highly customizable mapping
library(ggthemes)      # for more themes (including theme_map())
library(plotly)        # for the ggplotly() - basic interactivity
library(gganimate)     # for adding animation layers to ggplots
library(gifski)        # for creating the gif (don't need to load this library every time,but need it installed)
library(transformr)    # for "tweening" (gganimate)
library(shiny)         # for creating interactive apps
library(patchwork)     # for nicely combining ggplot2 graphs  
library(gt)            # for creating nice tables
library(rvest)         # for scraping data
library(robotstxt)     # for checking if you can scrape data
library(readxl)
library(scales)
theme_set(theme_minimal())
```

```{r data}
#Eviction data from the Eviction Lab
evictions_state <- read_csv("evictions_state.csv")
states_midpoint <- read_csv("state-midpoints.csv")

evictions_county <- read_csv("evictions_county.csv") %>% 
  separate(col = name, sep = " County", into = c("name", "drop")) %>% 
  select(-drop) %>% 
  mutate(name = str_replace_all(name, "St. ", "St "),
         name = str_replace_all(name, " Parish", ""))

evictions_county$name[evictions_county$name == "LaMoure"] <- "La Moure"
evictions_county$name[evictions_county$name == "DuPage"] <- "Du Page"
evictions_county$name[evictions_county$name == "O'Brien"] <- "OBrien"
evictions_county$name[evictions_county$name == "Ste. Genevieve"] <- "Ste Genevieve"

states_map <- map_data("state")
county_map <- map_data("county")
```

### The Data
Our data is from the [Eviction Lab](https://evictionlab.org/). On their website, they introduce themselves as "The Eviction Lab is a team of researchers, students, and website architects who believe that a stable, affordable home is central to human flourishing and economic mobility. Accordingly, understanding the sudden, traumatic loss of home through eviction is foundational to understanding poverty in America. Drawing on tens of millions of records, the Eviction Lab at Princeton University has published the first-ever dataset of evictions in America, going back to 2000." One of the main variables of interest is "Eviction Rate," which is a measure of the number of evictions per 100 renter homes. The data also includes demographic data on both state and county levels.

### Our Initial Research Questions
When we began thinking about the data, we had some questions right away about eviction and race. Is there a relationship between race and the eviction rate in the US? There is also rent information in the data, and we wanted to compare rent over each state and see what effect rent had on eviction rates. Some of the final questions we had were what are the states with the highest eviction filing rates, and if there are states where it is easier to evict people. 

### Some Preliminary Graphs
First, we wanted to explore the data with a plotly graph and mapped the 2016 eviction rates by state. After looking at that graph and discussing what we wanted to show with our data (Courtesy of Eviction Lab), we decided to make a shiny app.  We wanted an app that users could interact with to see eviction rates over time as well as interact with demographics and other data by state to see relationships between that data and eviction rates.  

```{r}
evict_rate_map <- evictions_state %>% 
  filter(year == 2016) %>% 
  mutate(name = str_to_lower(name)) %>% 
  ggplot() +
  geom_map(map = states_map,
           aes(map_id = name,
               fill = `eviction-rate`,
               text = paste0(name, paste0(":<br>", `eviction-rate`)))) +
  expand_limits(x = states_map$long, y = states_map$lat) + 
  theme_map() +
  labs(title = "",
       fill = "") +
  scale_fill_viridis_c() +
  theme(legend.background = element_blank()) +
  labs(title = "Eviction Rate by State in 2016")

  ggplotly(evict_rate_map, tooltip = "text")
```

To help us look at the data and see how it is related, we made a correlation plot with race and the eviction-related variables. Unfortunately, nothing jumped out at us. We found it interesting that the percent of white people in a state was negatively correlated with all of the eviction variables, but the percent of African American people is positively correlated, though these relationships are both weak.

```{r}
corrplot <- evictions_state[, c(13:24)]

corrplot <- corrplot %>% 
  rename(
    `% White` = `pct-white`,
    `% African American` = `pct-af-am` ,
    `% Hispanic` = `pct-hispanic`,
    `% American Indian `= `pct-am-ind`,
    `% Asian` = `pct-asian`,
    `% Pacific Islander` = `pct-nh-pi`,
    `% Multiple` = `pct-multiple`,
    `% Other` = `pct-other`,
    `Eviction Filings` = `eviction-filings`,
    `Evictions` = `evictions`,
    `Eviction Rate` = `eviction-rate`,
    `Eviction Filing Rate` = `eviction-filing-rate`              
  )

corrplot <- round(cor(corrplot, use = "pairwise.complete.obs"),2)

library(reshape2)
melted_cormat <- melt(corrplot)

  # Get upper triangle of the correlation matrix
  get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
  }
  upper_tri <- get_upper_tri(corrplot)

melted_cormat <- melt(upper_tri, na.rm = TRUE)
ggplot(data = melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 10, hjust = 1),
    plot.title.position = "plot")+
  labs(title = "Eviction Correlation Plot with Race Related Variables",
       y = "",
       x = "") +
 coord_fixed() 
```

When looking at a similar plot with rent, population data, and the relationship with the eviction variables, once again, nothing jumps out at us that is not already to be expected. There is a strong positive relationship between evictions, population, and renter-occupied households. We would expect this to be true since places with more people will have more evictions. Other variable relationships with Eviction Rate are all very weak, and we do not feel confident about a causal relationship.

```{r}
corrplot2 <- evictions_state[, c(5:12, 21:24)]

corrplot2 <- corrplot2 %>% 
  rename(
    `Population` = `population`,
    `Poverty Rate` = `poverty-rate` ,
    `Renter Occupied Households` = `renter-occupied-households`,
    `% Renter Occupied`= `pct-renter-occupied`,
    `Median Gross Rent` = `median-gross-rent`,
    `Median Household Income` = `median-household-income`,
    `Median Property Value` = `median-property-value`,
    `Rent Burden` = `rent-burden`,
    `Eviction Filings` = `eviction-filings`,
    `Evictions` = `evictions`,
    `Eviction Rate` = `eviction-rate`,
    `Eviction Filing Rate` = `eviction-filing-rate`              
  )

corrplot2 <- round(cor(corrplot2, use = "pairwise.complete.obs"),2)

library(reshape2)
melted_cormat <- melt(corrplot2)

# Get upper triangle of the correlation matrix
  get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
  }
upper_tri <- get_upper_tri(corrplot2)

melted_cormat <- melt(upper_tri, na.rm = TRUE)
ggplot(data = melted_cormat, aes(Var2, Var1, fill = value)) +
 geom_tile(color = "white") +
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 10, hjust = 1),
    plot.title.position = "plot")+
  labs(title = "Eviction Correlation Plot with Rent Related Variables",
       y = "",
       x = "") +
 coord_fixed() 
```

This graph, for example, overlays eviction rates on a map of the US that is colored by population. We can see that population does not substantially affect eviction rates as California and Texas do not have significantly high eviction rates compared to states like South Carolina and Delaware, which have much lower populations.

```{r}
options(scipen = 9)
evictions_state %>% 
  left_join(states_midpoint,
            by = c("name" = "location")) %>% 
  filter(year == 2016,
         name != "Hawaii") %>% 
  mutate(name = str_to_lower(name)) %>% 
  ggplot() +
  geom_map(map = states_map,
           aes(map_id = name,
               fill = population)) +
  geom_point(aes(x = lon, y = lat, size = `eviction-rate`),
             color = "tomato") +
  expand_limits(x = states_map$long, y = states_map$lat) + 
  theme_map() +
  labs(title = "",
       fill = "Population",
       size = "Eviction Rate") +
  scale_fill_viridis_c(labels = comma) +
  theme(legend.background = element_blank(),
        legend.position = "right") +
  labs(title = "Population and Eviction rates in 2016")
```

### Race and Evictions in Minnesota

Looking at the counties with the highest percentage of African Americans in Minnesota compared to the counties with the highest eviction rates shows that there is not much overlap in Minnesota. Hennepin County (the county with the highest percentage of African Americans) has a very average eviction rate compared to the other counties in Minnesota. Hennepin County does have the highest number of evictions in the state, and while the rate may be low, there is a gap in the data. We do not know if African American renters are evicted at higher rates because we do not have eviction data by race. So while the rate in the county with the highest number of evictions is low, it could be disproportionately African Americans who are evicted.

```{r}
mn_map <- county_map %>% filter(region == "minnesota") %>% select(-region) %>% dplyr::rename(region = subregion)

race <- evictions_county %>% 
  filter(`parent-location` == "Minnesota",
         year == 2016) %>% 
  mutate(lwr_name = str_to_lower(name)) %>% 
  ggplot(aes(fill = `pct-af-am`)) +
  geom_map(map = mn_map,
           aes(map_id = lwr_name)) +
  scale_fill_viridis_c(labels = comma) +
  expand_limits(x = mn_map$long, y = mn_map$lat) + 
  theme_map() +
  labs(title = "% African American",
       fill = "",
       caption = "Data from 2016") +
  theme(legend.background = element_blank(),
            legend.position = "right")

eviction_rate <- evictions_county %>% 
  filter(`parent-location` == "Minnesota",
         year == 2016) %>% 
  mutate(lwr_name = str_to_lower(name)) %>% 
  ggplot(aes(fill = `eviction-rate`)) +
  geom_map(map = mn_map,
           aes(map_id = lwr_name)) +
  scale_fill_viridis_c(labels = comma) +
  expand_limits(x = mn_map$long, y = mn_map$lat) + 
  theme_map() +
  labs(title = "Eviction Rate",
       fill = "",
       caption = "Data from 2016") +
  theme(legend.background = element_blank(),
            legend.position = "right")

race | eviction_rate

evictions <- evictions_county %>% 
  filter(`parent-location` == "Minnesota",
         year == 2016) %>% 
  mutate(lwr_name = str_to_lower(name)) %>% 
  ggplot(aes(fill = `evictions`)) +
  geom_map(map = mn_map,
           aes(map_id = lwr_name)) +
  scale_fill_viridis_c(labels = comma) +
  expand_limits(x = mn_map$long, y = mn_map$lat) + 
  theme_map() +
  labs(title = "Number of Evictions",
       fill = "",
       caption = "Data from 2016") +
  theme(legend.background = element_blank(),
            legend.position = "right")

race | evictions
```

### Our Shiny App
Click [Here](https://alexhopkins.shinyapps.io/US-Evictions/) to interact with the data yourself! Choose from various factors like race, median rent, and year to display graphs to see relationships between demographics and evictions. The first tab displays statewide data, and the second "by state" tab displays individual states and their county-level data. Both tabs have a check box that allows the user to overlay eviction rate data over states and counties. This allows for a comparison between the eviction rate and the demographics of those places.

### Final Remarks
We observe that areas with larger populations have higher eviction rates, which is pretty straightforward. More people means more renters, which means more chances for evictions. States like Nevada and Arizona had large eviction rates at first, but over time they have diminished, while states like South Carolina have seen considerable rises in their eviction rates. In 2008-09 the eviction crisis was high, mirroring the economic crisis. We hope that creating this app will make it easier to make insights about eviction data in the U.S. and produce easy-to-understand visualizations of the eviction data.

